---
title: "Write out files for open ended answers"
output: html_document
---

```{r}
library(tidyverse)
library(kableExtra)
```


# Study 1

Study 1 included open-ended answers for general comments on the survey. 

```{r}
# read wide version data set
exp1_wide <- read_csv("../exp_1/data/cleaned_wide.csv") 
```


```{r}
# make an overview table

# levels(as.factor(exp1_wide$comments))
table <- exp1_wide %>% 
  drop_na(comments) %>% 
  select(id, comments)

# Create the basic table
if (knitr::is_latex_output() || knitr::is_html_output()) {
  # For LaTeX or HTML: Use kbl, column_spec, and kable_paper
  table_output <- table %>%
    kbl(
      booktabs = TRUE, 
      longtable = TRUE,
      caption = "Comments of participants on Study 1"
    ) %>%
    kableExtra::kable_paper(full_width = FALSE) %>%
    kableExtra::column_spec(1) %>%
    kableExtra::column_spec(2, width = "40em")
} else {
  # For Word: Use flextable for better Word compatibility
  table_output <- flextable::flextable(table) %>%
    flextable::set_caption(caption = "Comments of participants on Study 1") %>%
    flextable::autofit() %>%
    flextable::width(j = 2, width = 4.5) # Adjust column widths as needed
}

# Display the table
# table_output
```


```{r}
# write out data
write_csv(table, "study-1_comments.csv")
```


# Study 2

Study 2 included open-ended answers for reasons for consensus rejection.

```{r}
# read wide version data set
exp2_wide <- read_csv("../exp_2/data/cleaned_wide.csv") 
```


```{r exp2-reasons-rejection}
# make an overview table

reasons_rejection <- c("electrons", "antibiotics", "continents", "sex", "lasers", 
               "orbit", "diamonds", "speed", "salt", "water")


table <- exp2_wide %>% 
  select(id, all_of(reasons_rejection)) %>% 
  pivot_longer(all_of(reasons_rejection), 
               names_to = "question",
               values_to = "answer") %>% 
  drop_na(answer) %>% 
  arrange(id)

# Create the basic table
if (knitr::is_latex_output() || knitr::is_html_output()) {
  # For LaTeX or HTML: Use kbl, column_spec, and kable_paper
  table_output <- table %>%
    kbl(
      booktabs = TRUE, 
      longtable = TRUE,
      caption = "Reasons for consensus rejection (raw answers)"
    ) %>%
    kableExtra::kable_paper(full_width = FALSE) %>%
    kableExtra::column_spec(1) %>%
    kableExtra::column_spec(2) %>%
    kableExtra::column_spec(3, width = "30em")
} else {
  # For Word: Use flextable for better Word compatibility
  table_output <- flextable::flextable(table) %>%
    flextable::set_caption(caption = "Reasons for consensus rejection (raw answers)") %>%
    flextable::autofit() %>%
    flextable::width(j = 3, width = 4.5) # Adjust column width for the third column
}

# Display the table
# table_output
```


```{r}
# write out data
write_csv(table, "study-2_reasons-for-rejection.csv")
```

# Study 3

Study 3 included open-ended answers for (i) reasons for consensus rejection, (ii) reasons for consensus acceptance if selected `other` and (iii) statements on how participants independently verified information (in case this applied). 

```{r}
# read wide version data set
exp3_wide <- read_csv("../exp_3/data/cleaned_wide.csv") 
```

## Reasons for consensus rejection

```{r exp3-reasons-rejection}
# store all variables
reasons_rejection <- c("electrons", "antibiotics", "continents", "sex", "lasers", 
               "orbit", "diamonds", "speed", "salt", "water")


table <- exp3_wide %>% 
  select(id, all_of(reasons_rejection)) %>% 
  pivot_longer(all_of(reasons_rejection), 
               names_to = "question",
               values_to = "answer") %>% 
  drop_na(answer) %>% 
  arrange(id)

# Create the basic table
if (knitr::is_latex_output() || knitr::is_html_output()) {
  # For LaTeX or HTML: Use kbl, column_spec, and kable_paper
  table_output <- table %>%
    kbl(
      booktabs = TRUE, 
      longtable = TRUE,
      caption = "Reasons for consensus rejection (raw answers)"
    ) %>%
    kableExtra::kable_paper(full_width = FALSE) %>%
    kableExtra::column_spec(1) %>%
    kableExtra::column_spec(2) %>%
    kableExtra::column_spec(3, width = "30em")
} else {
  # For Word: Use flextable for better Word compatibility
  table_output <- flextable::flextable(table) %>%
    flextable::set_caption(caption = "Reasons for consensus rejection (raw answers)") %>%
    flextable::autofit() %>%
    flextable::width(j = 3, width = 4.5) # Adjust column width for the third column
}

# Display the table
# table_output

```

```{r}
# write out data
write_csv(table, "study-3_reasons-for-rejection.csv")
```

## Reasons for consensus acceptance

```{r exp3-other-reasons-acceptance}
# store all variables
table <- exp3_wide %>% 
  select(id, other_reason_agreement) %>% 
  drop_na(other_reason_agreement) %>% 
  arrange(id)

# Create the basic table
if (knitr::is_latex_output() || knitr::is_html_output()) {
  # For LaTeX or HTML: Use kbl, column_spec, and kable_paper
  table_output <- table %>%
    kbl(
      booktabs = TRUE, 
      longtable = TRUE,
      caption = "Reasons for consensus acceptance (participants who answered 'other')"
    ) %>%
    kableExtra::kable_paper(full_width = FALSE) %>%
    kableExtra::column_spec(1) %>%
    kableExtra::column_spec(2, width = "30em")
} else {
  # For Word: Use flextable for better Word compatibility
  table_output <- flextable::flextable(table) %>%
    flextable::set_caption(caption = "Reasons for consensus acceptance (participants who answered 'other')") %>%
    flextable::autofit() %>%
    flextable::width(j = 2, width = 4.5) # Adjust column width for the second column
}

# Display the table
# table_output
```

```{r}
# write out data
write_csv(table, "study-3_reasons-for-acceptance-if-other.csv")
```

## How did participants independently verifiy?

```{r exp3-independent-verification}
# store all variables
table <- exp3_wide %>% 
  select(id, reason_followup) %>% 
  drop_na(reason_followup) %>% 
  arrange(id)

# Create the basic table
if (knitr::is_latex_output() || knitr::is_html_output()) {
  # For LaTeX or HTML: Use kbl, column_spec, and kable_paper
  table_output <- table %>%
    kbl(
      booktabs = TRUE, 
      longtable = TRUE,
      caption = "Answers from participants saying they trusted the scientific consensus because they verified independently on how they did so."
    ) %>%
    kableExtra::kable_paper(full_width = FALSE) %>%
    kableExtra::column_spec(1) %>%
    kableExtra::column_spec(2, width = "30em")
} else {
  # For Word: Use flextable for better Word compatibility
  table_output <- flextable::flextable(table) %>%
    flextable::set_caption(caption = "Answers from participants saying they trusted the scientific consensus because they verified independently on how they did so.") %>%
    flextable::autofit() %>%
    flextable::width(j = 2, width = 4.5) # Adjust column width for the second column
}

# Display the table
# table_output
```


```{r}
# write out data
write_csv(table, "study-3_reasons-if-independent-verification.csv")
```


# Study 4

Study 3 included open-ended answers for (i) reasons for consensus rejection, (ii) reasons for consensus acceptance if selected `other` and (iii) statements on how participants independently verified information (in case this applied). 

```{r}
# read data

# read the cleaned long version data set
exp4_long <- read_csv("../exp_4/data/cleaned_long.csv")

# read wide version data set
exp4_wide <- read_csv("../exp_4/data/cleaned_wide.csv") 

# read coded justifications
exp4_justifications <- read_csv("../exp_4/data/justifications_clean.csv")
```

## Reasons for consensus rejection

```{r exp4-reasons-rejection}
# store all variables
reasons_rejection <- unique(exp4_long$subject)


table <- exp4_wide %>% 
  select(id, all_of(reasons_rejection)) %>% 
  pivot_longer(all_of(reasons_rejection), 
               names_to = "question",
               values_to = "answer") %>% 
  drop_na(answer) %>% 
  arrange(id)

# Create the basic table
if (knitr::is_latex_output() || knitr::is_html_output()) {
  # For LaTeX or HTML: Use kbl, column_spec, and kable_paper
  table_output <- table %>%
    kbl(
      booktabs = TRUE, 
      longtable = TRUE,
      caption = "Reasons for consensus rejection (raw answers)"
    ) %>%
    kableExtra::kable_paper(full_width = FALSE) %>%
    kableExtra::column_spec(1) %>%
    kableExtra::column_spec(2) %>%
    kableExtra::column_spec(3, width = "30em")
} else {
  # For Word: Use flextable for better Word compatibility
  table_output <- flextable::flextable(table) %>%
    flextable::set_caption(caption = "Reasons for consensus rejection (raw answers)") %>%
    flextable::autofit() %>%
    flextable::width(j = 3, width = 4.5) # Adjust column width for the third column
}

# Display the table
# table_output
```

```{r}
# write out data
write_csv(table, "study-4_reasons-for-rejection.csv")
```

## Reasons for consensus acceptance

```{r exp4-other-reasons-acceptance}
# store all variables
table <- exp4_wide %>% 
  select(id, other_reason_agreement) %>% 
  drop_na(other_reason_agreement) %>% 
  arrange(id)

# Create the basic table
if (knitr::is_latex_output() || knitr::is_html_output()) {
  # For LaTeX or HTML: Use kbl, column_spec, and kable_paper
  table_output <- table %>%
    kbl(
      booktabs = TRUE, 
      longtable = TRUE,
      caption = "Reasons for consensus acceptance (participants who answered 'other')"
    ) %>%
    kableExtra::kable_paper(full_width = FALSE) %>%
    kableExtra::column_spec(1) %>%
    kableExtra::column_spec(2, width = "30em")
} else {
  # For Word: Use flextable for better Word compatibility
  table_output <- flextable::flextable(table) %>%
    flextable::set_caption(caption = "Reasons for consensus acceptance (participants who answered 'other')") %>%
    flextable::autofit() %>%
    flextable::width(j = 2, width = 4.5) # Adjust column width for the second column
}

# Display the table
# table_output
```

```{r}
# write out data
write_csv(table, "study-4_reasons-for-acceptance-if-other.csv")
```

## How did participants independently verifiy?

```{r exp4-independent-verification}
# store all variables
table <- exp4_wide %>% 
  select(id, reason_followup) %>% 
  drop_na(reason_followup) %>% 
  arrange(id)

# Create the basic table
if (knitr::is_latex_output() || knitr::is_html_output()) {
  # For LaTeX or HTML: Use kbl, column_spec, and kable_paper
  table_output <- table %>%
    kbl(
      booktabs = TRUE, 
      longtable = TRUE,
      caption = "Answers from participants saying they trusted the scientific consensus because they verified independently on how they did so."
    ) %>%
    kableExtra::kable_paper(full_width = FALSE) %>%
    kableExtra::column_spec(1) %>%
    kableExtra::column_spec(2, width = "30em")
} else {
  # For Word: Use flextable for better Word compatibility
  table_output <- flextable::flextable(table) %>%
    flextable::set_caption(caption = "Answers from participants saying they trusted the scientific consensus because they verified independently on how they did so.") %>%
    flextable::autofit() %>%
    flextable::width(j = 2, width = 4.5) # Adjust column width for the second column
}

# Display the table
# table_output
```

```{r}
# write out data
write_csv(table, "study-4_reasons-if-independent-verification.csv")
```








